Forecasting short-term hourly Emergency Department arrivals

### 2022-01-13 paper/paper.Rmd
add points instead of labels in the plot of time versus accracy

abstract: |
   An accurate forecast of Emergency Department (ED) arrivals by an hour of the day is critical to meet patients’ demand. It enables planners to match ED staff to the number of arrivals, redeploy staff, and reconfigure units. This can have many advantages for healthcare staff and the quality of care delivered to patients. In this study, we develop an innovative model based on Generalised Additive Models and an advanced dynamic model based on exponential smoothing to generate an hourly probabilistic forecast of ED arrivals. We compare the forecast accuracy of these models against appropriate benchmarks, including TBATS, Poisson Regression, Prophet, and simple empirical distribution. We use Root Mean Squared Error (RMSE) to examine the point forecast accuracy and assess the forecast distribution accuracy using Quantile Bias, PinBall Score and Pinball Skill Score. Our results indicate that the proposed models outperform their benchmarks for point and probabilistic forecasts. Our developed models can also be generalised to forecast hourly arrivals in other services such as hospitals, ambulances, or clinical desk services.

@@

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, fig.pos = 'H', message = FALSE, warning = FALSE)
# Make sure you have the latest versions of rmarkdown and rticle
library(tidyverse)
library(lubridate)
library(bookdown)
library(patchwork)
library(tsibble)
library(kableExtra)
library(ggplot2)
library(RColorBrewer)
library(ggridges)
library(hrbrthemes)
library(ggthemes)
library(scales)
library(ggtext)
library(extrafont)
library(feasts)
```

@@

Forecasting Emergency Department (ED) arrivals is critical for informing staffing and scheduling decisions to meet the needs of patients. Accurate ED demand forecasts contribute to a better decision making process regarding resources allocation and staffing. This is one of the best ways to optimise resources utilisation and minimise related costs. An accurate forecast of patient arrivals is crucial in ED services to depict various courses of action that can result in massive savings in terms of patient lives. Inability to match the staff with the demand might result in patients overcrowding the system, which is a severe problem that causes challenges for the patient flow [@derlet2002overcrowding]. Also, it is related to increasing length of stay [@muhammet2015forecasting], low patient satisfaction, unexpected return visits to services, increased health care costs, inaccuracy in electronic medical records and other [@rostami2020anticipating].

Accurate forecasting of arrivals by the hour of the day enables planners to match staff to meet anticipated patients, reconfigure units and redeploy staff. This has many advantages for both patients, staff and the quality of provided services. Hourly forecasts are required to inform the short-term operational planning for the current and the upcoming shifts of the day. This involves the short-term decision making related to the execution of the delivery process in ED. The combination of an hourly arrival forecast, current staff, being occupied, resource availability and waiting times at ED provide information on the state of the unscheduled care system across the service. Having this complete picture enables the delivery managers to focus on the areas that require intervention to allow the most effective delivery of the service to the patients. However, compared with lower frequency time series forecasting such as monthly, quarterly and yearly, hourly forecasts are challenging because the noise caused by random variation may overshadow any pattern in the time series. Hourly time series generally exhibit multiple seasonal cycles of different lengths: hourly, daily, weekly, and yearly. They may also express nonstationarity, and their profile may change over time. Therefore, an appropriate forecasting model should consider these features to accurately predict hourly demand admissions.

There are few studies that look at forecasting hourly arrivals in ED and other hospital services using historical time series data and/or predictors such as patient characteristics, weather, holidays and public events. These studies use multiple approaches, including Exponential Smoothing [@SvetunkovAdam], Autoregressive Integrated Moving Average (ARIMA) [@hyndman2021forecasting], Autoregressive Conditional Heteroskedasticity (ARCH) [@bollerslev1994arch], Vector Autoregressive model [@lutkepohl2013vector],  TBATS [@de2011forecasting] and Artificial Neural Networks [@hyndman2021forecasting]. However, most of these studies are limited to only predicting future arrivals as a point forecast (a single number), which does not quantify any uncertainty associated with the number of future arrivals. There are few studies that report uncertainty by presenting prediction intervals, but there is no study generating and evaluating the entire forecast distribution of arrivals. Reporting the uncertainty via the forecast distribution is potentially valuable in this setting because the consequences of inadequate staffing are asymmetric. This asymmetry arises because it is preferable to incur a small opportunity cost associated with utilised staff rather than compromise service levels if staff levels are insufficient. Probabilistic forecasts inform decision-makers about exposure to these risks and potentially enable those risks to be managed more efficiently.

Furthermore, if the impact of under- and overstaffing can be quantified, probabilistic forecasts allow `optimal' decisions that balance the cost associated with under- and overstaffing to be calculated. Therefore, in this paper, in addition to generating point forecasts, we also produce and evaluate density forecasts of hourly ED arrivals, comparing several methods for this task. Another drawback of existing studies is that the datasets used are relatively small (e.g. time period of 1-2 years), making it challenging to report the forecast accuracy using robust approaches such as time series cross-validation. Such results might not be generalisable. Additionally, most of the forecasting methods used in these publications do not consider the full extent of the multiple seasonality of hourly ED arrivals. Moreover, all previous publications referenced in this paper are not fully reproducible as underlying data and code are not available.

@@

1. We produce probabilistic forecasts, in addition to the point estimation, quantifying uncertainties in future hospital admission, and comparing different forecasting methods using a suite of well-established evaluation metrics;

2. We develop an advanced dynamic model to forecast ED arrivals based on iETS [@Svetunkov2019a] and ETSX models with a modification for multiple frequencies, which produced highly-accurate point forecasts;

3. We develop a novel model to produce a probabilistic forecast of ED arrivals based on Generalised Additive Models for Location Scale and Shape, which accounts for i) the bounded and non-Gaussian distribution of arrivals, ii) multiple seasonalities, weather and holiday effects, and  iii) variation in forecast uncertainty;

4. We benchmark the accuracy of our model against appropriate models used when multiple seasonality is present, i.e. Prophet, TBATS, Poisson Regression, Exponential Smoothing State Space model (ETS) and the simple empirical distribution of the arrivals;

5. We provide data and code enabling reproduction and refinement of the proposed approach and benchmarks. The proposed approach could also be generalised to forecast hourly requirements in other services, such as the number of incidents or call volumes in clinical desk services.

The rest of the paper is organised as follows: In section \@ref(lit), we provide a brief overview of hourly ED arrival forecasting; In Section \@ref(data), we present the hourly time series of an ED arrival and use various plots to highlight important patterns. In section \@ref(model), we describe the modelling approach and benchmark methods. We then discuss the performance evaluation metrics in section \@ref(accuracy); in section \@ref(result), we present and discuss our results. Finally, we summarise our findings and present ideas for future research in section \@ref(conclusion).

@@

Some studies employ models to forecast admissions and arrivals in Emergency Department. The time granularity considered by these studies spans from hourly to yearly. However, given the focus of the paper, we only discuss hourly ED forecasting.

@hertzum2017forecasting used linear regression, ARIMA, and Naïve to investigate whether accurate hourly accident and emergency department patient arrivals and occupancy forecasts can be generated using calendar variables. @hertzum2017forecasting study found that patient arrival variation is larger across the hours of the day than across the days of the week and the months of the year. In terms of the hour of the day, patient arrivals peaked around noon. For days of the week, Monday is the busiest day, while weekends are the quietest ones. July-August are the months with the highest number of patient arrivals, while January and February have the lowest numbers. They indicate that regression and ARIMA models performed similarly in modelling patient arrivals, while ARIMA outperformed regression models in modelling accident and emergency department occupancy.

@choudhury2020forecasting used ARIMA, Holt-Winters, TBATS, and neural network methods to forecast hourly accident and emergency department arrivals. ARIMA model was selected as the best fit model. Authors claimed that ARIMA provided high and acceptable hourly ED forecasting accuracy, even outperforming TBATS. @Cheng2021 developed an ARIMA model for ED occupancy with a seasonal component and exogenous variables, which outperformed a rolling-average benchmark. They also produce prediction intervals, a form of the probabilistic forecast, which were found to be well-calibrated, a necessary property for such forecasts.

@morzuch2006forecasting used the Unobserved Components Model (UCM), in which each component of the time series is separately modelled as stochastic. Double-seasonal exponential smoothing and standard Holt-Winters were used to forecast ED arrival for a horizon of 168 hours. The hourly data collected from an ED in Pennsylvania showed no trend and two seasonal cycles: a within-day and a within-week seasonal cycles. The double seasonal model recorded lower RMSEs for all the 168-hour horizons, which was expected due to the strong hourly seasonality of the time series.

@mccarthy2008challenge employed a Poisson log-linear regression model, including independent variables such as temporal factors (e.g., hour-of-day, day-of-week, type-of-day, season, and calendar year), patient characteristics (i.e., age, gender, insurance status, triage level, mode of arrival, and ambulance diversion status) and climatic factors (i.e., temperature and precipitation) to model patient demand for ED services. The authors produced probabilistic predictions in the form of $50\%$ and $90\%$ prediction intervals for the number of hourly arrivals. Hourly data of ED arrivals in the 1-year study period was modelled and analysed, and it was suggested that the model could be used for forecasting. However, model evaluation was performed in-sample on only one year of data, so it is unclear how this approach would perform in a forecasting setting or compare to simpler approaches. However, the length of the time series in this study was very short (only one year), which did not allow for a rigorous out-of-sample evaluation.

@schweigler2009forecasting investigated whether time series methods could accurately generate short-term forecasts of ED bed occupancy. A year-long dataset of hourly ED bed occupancy was collected from three facilities. The authors implemented an hourly historical average model, SARIMA model, and sinusoidal model with autocorrelated error for each facility. The historical average model was based on the mean occupancy for each site, for each hour of the day, while the sinusoidal model was based on four parameters: an AR term, a sine coefficient, a cosine coefficient and an intercept. They evaluated the forecast accuracy of four and twelve hours forecast horizons using RMSE. They found that both SARIMA and the sinusoidal models outperformed the historical average (for example, at site 2, the two models improved by $33\%$ the 12-hour forecasts generated by the historical average).

@kim2014predicting compared different univariate and multivariate time series forecasting techniques to predict patient volume for a Hospital Medicine programme. The study evaluated linear regression, exponential smoothing, ARIMA, SARIMA, Generalized Autoregressive Conditional Heteroskedasticity (GARCH) and  Vector Autoregressive (VAR) models to forecast for 4 and 24 hours ahead. They used Mean Absolute Percentage Error (MAPE) to report the forecast accuracy. The authors found that the ARIMA outperformed all the other models.

@@

@asheim2019real developed a Poisson time-series regression model with continuous day-of-week and week-of-year effects to implement a real-time system that could forecast ED arrivals on 1, 2, 3 hours ahead. Measuring the accuracy using the MAPE, @asheim2019real noticed that significant improvement happened when the time of notification was incorporated into the model, especially in the one-hour horizon.

@cheng2021forecasting used one year of ED visits time series to evaluate the Rolling Average, SARIMAX, ARIMA, VAR and Holt-Winter to forecast ED occupancy up to 4-hours ahead. The forecast accuracy is evaluated using Mean Squared Error (MSE), Mean Absolute Error (MAE) and MAPE for point forecast and coverage for prediction intervals of $80\%$ and $95\%$. They show that SARIMAX provides a more accurate forecast of hourly ED occupancy.

According to the studies mentioned above, it can be said that they have shown complications in forecasting hourly patient accident and emergency department visits, and the application of forecasting hourly patient visits is not well established. Some of the studies claimed that the accuracy of forecasting models on hourly accident and emergency department data is low [@boyle2012predicting; @hertzum2017forecasting], while others mentioned that the accuracy of ED hourly forecast is at the acceptable level [@choudhury2020forecasting; @mccarthy2008challenge;  @schweigler2009forecasting].

There are a few limitations in the literature which encourage us to undertake this research and examine different forecasting approaches:

(i) Current approaches to forecast hourly ED arrivals do not fully consider the feature of data such as multiple seasonal cycles and changing profile over time;
(ii) Almost all research studies produce point forecasts and, at best, report prediction intervals. There is a lack of studies presenting the entire forecast distribution of hourly ED arrivals that better represent the uncertainty of future arrivals, providing a holistic picture of future demand for a planner;
(iii) most studies are not reproducible, as it is almost impossible to reapply the approaches without the help of the authors of those papers;
(iv) studies are limited in terms of the length of historical data used for training purposes and forecast performance evaluation and
(v) some studies in this area lack a rigorous experimental design, i.e. they do not use benchmark methods or report forecast accuracy.

@@

```{r tab-results, out.width= "50%"}
results_table <- read_rds("results_table.rds")
knitr::kable(results_table, booktabs = T, 
             linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("hold_position"), font_size =11)
```

>>>>

abstract: |
   An accurate forecast of Emergency Department (ED) arrivals by hour of the day is critical to meet patients’ demand It will enable planners to match ED staff to number of arrivals, redeploy staff and reconfigure units, if required. This can have many advantages for healthcare staff and the quality of care delivered to patients. In this study, we develop an innovative model based on Generalised Additive Models and an advanced dynamic model based on exponential smoothing, to generate hourly probabilistic forecast of ED arrivals. We compare the forecast accuracy of these models against appropriate benchmarks including TBATS, Poisson Regression, Prophet, and a simple empirical distribution. We use Root Mean Squared Error (RMSE) to examine the point forecast accuracy and the forecast distribution accuracy is assessed using Quantile Bias, PinBall Score and Pinball Skill Score. Our results indicate that the proposed models outperform their benchmarks for both point and probabilistic forecasts. Our developed models can also be generalised to forecast hourly arrivals in other services such as hospitals, ambulance, or clinical desk services.

@@

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, fig.pos = 'H', message = FALSE, warning = FALSE)
# Make sure you have the latest versions of rmarkdown and rticle
library(tidyverse)
library(lubridate)
library(bookdown)
library(patchwork)
library(tsibble)
library(kableExtra)
library(ggplot2)
library(RColorBrewer)
library(ggridges)
library(hrbrthemes)
library(ggthemes)
library(scales)
library(ggtext)
library(extrafont)
library(feasts)
library(data.table)
```

@@

Forecasting Emergency Department (ED) arrivals is critical for informing staffing and scheduling decisions to meet the needs of patients. An accurate ED demand forecasts contribute to a better decision making process regarding the resources allocation and staffing. This is one of the best ways to optimize resources utilisation and minimise related costs. 
<!-- State of the art practice to optimise personnel scheduling follows the general approach originally presented by @vile2016time, which recommends that the following steps be taken to roster employees: (i) forecast demand; (ii) convert demand forecasts into staffing requirements; (iii) schedule shifts optimally; and (iv) assign employees to shifts.  -->
An accurate forecast of patient arrivals is crucial in ED services to depict various courses of action that can result in massive savings in terms of patient lives. Inability to match the staff with the demand might result in patients overcrowding the system which is a serious problem that causes challenges for the patient flow [@derlet2002overcrowding]. Also, it is related with increasing length of stay [@muhammet2015forecasting], low patient satisfaction, unexpected return visits to services, increased health care costs, inaccuracy in electronic medical records and other [@rostami2020anticipating].
<!-- In this respect, forecasts of daily or monthly arrivals are likely easier but target decisions about staff allocation and the like, not the ongoing scheduling and rescheduling of how the available resources are divided among the patients in need of emergency services -->

An accurate forecasting of arrivals by hour of the day enables planners to match staff to meet anticipated patients, reconfigure units and redeploy staff. This has many advantages for both patients, staff and the quality of provided services. Hourly forecasts are required to inform the short-term operational planning for the current and the upcoming shifts of the day. This involves the short-term decision making related to the execution of the delivery process in ED.
<!-- for various health care services such as Ambulatory, Emergency, Surgical, Inpatient, Home and Residential.  -->
The combination of an hourly arrival forecast, current staff being occupied, resource availability and waiting times at ED, provide information on the state of the unscheduled care system across the service. Having this full picture enables the delivery managers to focus on the areas that require intervention to enable the most effective delivery of the service to the patients. However, in comparison with lower frequency time series forecasting such as monthly, quarterly and yearly, hourly forecasts are challenging because the noise caused by random variation may overshadow any pattern in the time series. Hourly time series generally exhibit multiple seasonal cycles of different lengths such as hourly, daily, weekly and yearly. They may also express nonstationarity and their profile may change over time. Therefore, an appropriate forecasting model should take these features into account, to accurately predict hourly demand admissions.
<!-- Moreover, there are often unplanned events, such as walk-in patients, extended consultation times, during  a day or a week. Staff capacities may be adjusted based on the predicted demand fluctuations by using part-time, on-call nurses, staff overtime and voluntary absenteeism. That may also require to adjust the scheduling. -->

There are few studies that look at forecasting hourly arrivals in ED and other hospital services using historical time series data and/or predictors such as patient characteristics, weather, holidays and public events. These studies use multiple approaches including Exponential Smoothing [@SvetunkovAdam], Autoregressive Integrated Moving Average (ARIMA) [@hyndman2021forecasting], Autoregressive Conditional Heteroskedasticity (ARCH) [@bollerslev1994arch], Vector Autoregressive model [@lutkepohl2013vector],  TBATS [@de2011forecasting] and Artificial Neural Networks [@hyndman2021forecasting]. However, majority of these studies are limited to only predict future arrivals as a point forecast (a single number), which does not quantify any uncertainty associated with the number of future arrivals. There are few studies that report uncertainty by presenting prediction intervals, no study generates and evaluates the entire forecast distribution of arrivals. Reporting the uncertainty via the forecast distribution is potentially valuable in this setting because the consequences of imperfect staffing are asymmetric. This asymmetry arises because it is preferable to incur a small opportunity cost associated under utilised staff rather than compromise service levels, if staff levels are insufficient. Probabilistic forecasts inform decision-makers about exposure to these risks and potentially enable those risks to be managed more efficiently. Furthermore, if the impact of under and over staffing can be quantified, probabilistic forecasts allow `optimal' decisions that balance the cost associated with under and over staffing to be calculated. Therefore, in this paper, in addition to generating point forecasts, we also produce and evaluate density forecasts of hourly ED arrivals, comparing several methods for this task. Another drawback of existing studies is that the datasets used are relatively small (e.g. time period of 1-2 year), which make it challenging to report the forecast accuracy using robust approaches such as time series cross validation and such results might not be generalisable. Additionally, most of forecasting methods used in these publications does not consider the full extend of multiple seasonality exhibited by hourly ED arrivals. Moreover, all previous publications referenced in this paper are not fully reproducible as underlying data and code are not available.

@@

1. We produce probabilistic forecasts, in addition to the point estimation, which quantify uncertainties in future hospital admission, and compare different forecasting methods using a suite of established evaluation metrics;

2. We develop an advanced dynamic model to forecasts ED arrivals based on iETS [@Svetunkov2019a] and ETSX models with a modification for multiple frequencies which produced highly-accurate point forecasts;

3. We develop a novel model to produce probabilistic forecast of ED arrivals based on Generalised Additive Models for Location Scale and Shape, which accounts for i) the bounded and non-Gaussian distribution of arrivals, ii) multiple seasonalities, weather and holiday effects, and  iii) variation in forecast uncertainty;

4. We benchmark the accuracy of our model against appropriate models used when multiple seasonality is present, i.e. Prophet, TBATS, Poisson Regression, Exponential Smoothing State Space model (ETS) and  a simple empirical distribution of the arrivals;

5. We provide data and code enabling reproduction and refinement of the proposed approach and benchmarks. The proposed approach could also be generalised to forecast hourly requirements in other services such as the number of incidents or call volumes in clinical desk services.

The rest of the paper is organised as following: In section \@ref(lit), we provide a brief overview of hourly ED arrival forecasting; In Section \@ref(data), we present the hourly time series of an ED arrival and use various plots to highlight important patterns. In section \@ref(model), we describe the modelling approach and benchmark methods. We then discuss the performance evaluation metrics in section \@ref(accuracy); in section \@ref(result), we present and discuss our results. Finally, we summarise our findings and present ideas for future research in section \@ref(conclusion).

@@

There are some studies that employ models to forecast admissions and arrivals in Emergency Department. The time granularity considered by these studies spans from hourly to yearly. However, given the focus of the paper, we only discuss the hourly ED forecasting.

Linear regression, ARIMA, and Naïve were used by @hertzum2017forecasting to investigate whether accurate hourly accident and emergency department patient arrivals and occupancy forecasts can be generated using calendar variables. @hertzum2017forecasting study found that patient arrival variation is larger across the hours of the day than across the days of the week and the months of the year. In terms of hour of the day, patient arrivals peaked around noon. For days of the week, Monday is the busiest day while weekends are the quietest ones. July-August are the months with the highest numbers of patient arrivals, while January and February are the months with the lowest numbers. They indicate that regression and ARIMA models performed similarly in modelling patient arrivals, while ARIMA outperformed regression models in modelling accident and emergency department occupancy.
<!-- Another interesting point is that the accuracy of accident and emergency department forecasting models decrease with the increasing forecast interval. Lastly, the accuracy of the forecasting model may possibly be increased with additional information added to the model. -->

@choudhury2020forecasting used ARIMA, Holt-Winters, TBATS, and neutral network methods to forecast hourly accident and emergency department arrivals. ARIMA model was selected as the best fit model. Authors claimed that ARIMA has provided high and acceptable hourly ED forecasting accuracy, even outperforming TBATS. @Cheng2021 develop an ARIMA model for ED occupancy with a seasonal component and exogenous variables, which is found to outperform a rolling-average benchmark. They also produce prediction intervals, a form of probabilistic forecast, which are found to be well calibrated, a necessary property for such forecasts.

@morzuch2006forecasting used the Unobserved Components Model (UCM), in which each component of the time series is separately modelled as stochastic. Double-seasonal exponential smoothing and standard Holt-Winters were used to forecast ED arrival for an horizon of 168 hours. The hourly data collected from an ED in Pennsylvania showed no trend, and two seasonal cycles: a within-day and a within-week seasonal cycles. The double seasonal model recorded lower RMSEs for all the 168-hour horizons, which was expected due to the strong hourly seasonality of the time series.

@mccarthy2008challenge employed a Poisson log-linear regression model, including independent variables such as temporal factors (e.g., hour-of-day, day- of-week, type-of-day, season, and calendar year), patient characteristics (i.e., age, gender, insurance status, triage level, mode of arrival, and ambulance diversion status) and climatic factors (i.e., temperature and precipitation) to model patient demand for ED services. The authors produced probabilistic predictions in the form of $50\%$ and $90\%$ prediction intervals for the number of hourly arrivals. Hourly data of ED arrivals in the 1-year study period was modelled and analysed, and it is suggested that the model could be used for forecasting, however, model evaluation was performed in-sample on only one year of data, so unclear how this approach would perform in a forecasting setting or compare to simpler approaches. However, the length of time series is very short (only one year), which does not allow for a rigorous out-of-sample evaluation.

@schweigler2009forecasting conducted an investigation on whether time series methods could accurately generate short-term forecasts of ED bed occupancy. A year-long dataset of hourly ED bed occupancy was collected from three facilities. For each facility, the authors implemented an hourly historical average model, SARIMA model and sinusoidal model with autocorrelated error. The historical average model was based on the mean occupancy for each site, for each hour of the day; while the sinusoidal model was based on 4 parameters: an AR term, a sine coefficient, a cosine coefficient and an intercept. They evaluated the forecast accuracy of four and twelve hours forecast horizon using RMSE and they found that both SARIMA and the sinusoidal models outperformed the historical average (for example, at site 2, the two models improved by $33\%$ the 12-hour forecasts generated by historical average).

@kim2014predicting compared different univariate and multivariate time series forecasting techniques to forecast patient volume for a Hospital Medicine programme. The study evaluated linear regression, exponential smoothing, ARIMA, SARIMA, Generalized Autoregressive Conditional Heteroskedasticity (GARCH) and  Vector Autoregressive (VAR) models to forecast for 4 hours and 24 hours ahead. They used Mean Absolute Percentage Error (MAPE) to report the forecast accuracy. ARIMA model outperformed all the other models.

@@

@asheim2019real developed a Poisson time-series regression model with continuous day-of-week and week-of-year effects to implement a real-time system that could forecast ED arrivals on 1, 2, 3 hours ahead. Measuring the accuracy using the MAPE, @asheim2019real noticed that great improvement happened when time of notification was incorporated into the model, especially in one-hour horizon.

@cheng2021forecasting used one year of ED visits time series to evaluate the performance Rolling Average, SARIMAX, ARIMA, VAR and Holt-Winter to forecast ED occupancy up to 4-hours ahead. The forecast accuracy is evaluated using Mean Squared Error (MSE), Mean Absolute Error (MAE) and MAPE for point forecast and coverage for prediction intervals of $80\%$ and $95\%$. They show that SARIMAX provides more accurately forecast of hourly ED occupancy.

According to the studies mentioned above, it can be said that they have shown complications in forecasting hourly patient accident and emergency department visits and the application of forecasting hourly patients visits is not well established. Some of the studies claimed that the accuracy of forecasting models on hourly accident and emergency department data forecasting model is low compared to the higher forecasting intervals, like daily [@boyle2012predicting; @hertzum2017forecasting]. While, others mentioned that the accuracy of ED hourly forecast is at the acceptable level [@choudhury2020forecasting; @mccarthy2008challenge;  @schweigler2009forecasting].

There are few limitations in the literature which encourage us to undertake this research and examine different forecasting approaches. These limitations are summarised as follows:

(i) Current approaches to forecast hourly ED arrivals do not fully consider the feature of data such as multiple seasonal cycles and changing profile over time;
(ii) Almost all research studies produce point forecasts and at best report prediction intervals. There is a lack of studies presenting the entire forecast distribution of hourly ED arrivals that better represent uncertainty of future arrivals, providing a holistic picture of future demand for a planner;
(iii) most studies are not reproducible, as it is almost impossible to reapply the approaches without the help of the authors of those papers;
(iv) studies are limited in terms of the length of historical data used for training purposes and forecast performance evaluation and
(v) some studies in this area lack a rigorous experimental design, i.e. there is no benchmark method or forecast accuracy is not reported.

@@

```{r tab-results, out.width= "50%", cache=FALSE}
results_table <- read_rds("results_table.rds")
knitr::kable(results_table, booktabs = T, 
             linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("hold_position"), font_size =11)
```
